Prediction of PIK3CA mutation with gene expression

### 2020-04-23 plosOne/plosOne.Rmd
prevalance of pik3ca

## Modeling

To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected. 5000 out of 20502 genes were included in the modeling process. Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package. [@kuhnRecipesPreprocessingTools2020] Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}. Lambda is  penalty scaling parameter and alpha is mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$). [@friedmanRegularizationPathsGeneralized2010] Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR). The modelin process was done with tidymodels rpackage.  [@kuhnTidymodelsEasilyInstall2020]

@@

@@

Figure 2 shows top 30 important predictors. IGF1R mRNA expression was the strongest negative predictor and PTEN was the strongest positive predictor. Both genes and PIK3CA are key players in tyrosin kanase pathway. The cancer type was important predictors. Some cancer types including uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are strongest predictors.

>>>>

10845 cases were available both PIK3CA mutation and mRNAD expression data. 5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at modelinf process method. 33 cancer type dummy variables were included in predictor variables.

The target variable was PIK3CA mutation status. The status of  PIK3CA was considered as positive when the case has following PIK3CA variants which is the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y. The therascreen PIK3CA RGQ PCR Kit was approved by U.S. Food and Drug adminidtration for PIK3CA inhibitor. We splited the three quarters of dataset for the trainset and one quarter for testset.

## Modeling process

To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected.  Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package. [@kuhnRecipesPreprocessingTools2020] Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}. Lambda is  penalty scaling parameter and alpha is mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$). [@friedmanRegularizationPathsGeneralized2010] Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR). The modelin process was done with tidymodels rpackage.  [@kuhnTidymodelsEasilyInstall2020]

@@

## Dataset summary 
- PIK3CA prevalence
-

@@

Figure 2 shows top 30 important predictors. The coefficient is the parameter of the predictor which represent the effect of the predictor on prediction. IGF1R mRNA expression was the strongest negative predictor and PTEN was the strongest positive predictor. Both genes and PIK3CA are key players in tyrosin kanase pathway. The cancer type was important predictors. Some cancer types including uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are strongest predictors.

### 2020-04-23 plosOne/plosOne.tex
prevalance of pik3ca

\hypertarget{modeling}{%
\subsection{Modeling}\label{modeling}}

To narrow down potential predictors, Genes with a large the median
absolute deviation (more than third-quartiles) were selected. 5000 out
of 20502 genes were included in the modeling process. Yeo-Johnson
transformation was done to correct skewness. Centering and scaling were
done. All preprocessing was done using recipe r package. {[}5{]}
Penalized logistic regression was applied to prediction modeling.
10-fold cross-validation with targe variable stratification was done
over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda is penalty scaling parameter
and alpha is mixing parameter of penalty function
(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
Model performance was evealuated with the area under receiver operating
characteristic (ROC) and area under the ROC (AUROC) and precision recall
(PR) and area under curve (AUPR). The modelin process was done with
tidymodels rpackage. {[}6{]}

@@

@@

In this study, the prevalence of PIK3CA mutation was 11\%. The low
prevalence of PIK3CA results in imbalaced dataset which makes the
prediction difficult. The AUROC is optimistic in terms of performance.
The AUPR is more informative than AUROC on imbalaced datasets. {[}7{]}

@@

\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
6. Kuhn M, Wickham H. Tidymodels: Easily install and load the
'tidymodels' packages. 2020.

\leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
7. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
Datasets. PLOS ONE. Public Library of Science; 10: e0118432.
doi:\href{https://doi.org/10.1371/journal.pone.0118432}{10.1371/journal.pone.0118432}

>>>>

10845 cases were available both PIK3CA mutation and mRNAD expression
data. 5128 out of 20502 genes were included in the modeling process
after filtering with median absolute deviation as described at modelinf
process method. 33 cancer type dummy variables were included in
predictor variables.

The target variable was PIK3CA mutation status. The status of PIK3CA was
considered as positive when the case has following PIK3CA variants which
is the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R,
E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
The therascreen PIK3CA RGQ PCR Kit was approved by U.S. Food and Drug
adminidtration for PIK3CA inhibitor. We splited the three quarters of
dataset for the trainset and one quarter for testset.

\hypertarget{modeling-process}{%
\subsection{Modeling process}\label{modeling-process}}

To narrow down potential predictors, Genes with a large the median
absolute deviation (more than third-quartiles) were selected.
Yeo-Johnson transformation was done to correct skewness. Centering and
scaling were done. All preprocessing was done using recipe r package.
{[}5{]} Penalized logistic regression was applied to prediction
modeling. 10-fold cross-validation with targe variable stratification
was done over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda is penalty scaling parameter
and alpha is mixing parameter of penalty function
(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
{[}6{]} Model performance was evealuated with the area under receiver
operating characteristic (ROC) and area under the ROC (AUROC) and
precision recall (PR) and area under curve (AUPR). The modelin process
was done with tidymodels rpackage. {[}7{]}

@@

\hypertarget{dataset-summary}{%
\subsection{Dataset summary}\label{dataset-summary}}

\begin{itemize}
\item
  PIK3CA prevalence
\item
\end{itemize}

@@

In this study, the prevalence of PIK3CA mutation was 11\%. The low
prevalence of PIK3CA results in imbalaced dataset which makes the
prediction difficult. The AUROC is optimistic in terms of performance.
The AUPR is more informative than AUROC on imbalaced datasets. {[}8{]}

@@

\leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
6. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
1--22.

\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
7. Kuhn M, Wickham H. Tidymodels: Easily install and load the
'tidymodels' packages. 2020.

\leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
8. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
Datasets. PLOS ONE. Public Library of Science; 10: e0118432.
doi:\href{https://doi.org/10.1371/journal.pone.0118432}{10.1371/journal.pone.0118432}
